The Bot on Speaking Terms: The Effects of Conversation Architecture on Perceptions of Conversational Agents

### 2023-02-15 paper.tex
updated perception and architecture taxonomy based on AK feedback

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Anon}

@@

\multirow{3}{*}{\parbox{0.16\textwidth}{Perception of Interaction with Agent}}
    & Usability & accuracy, ease of use, efficiency, helpfulness & 23 

@@

    & Engagement & enjoyment, annoyance, desirable, intention to use  & 28

@@

    & Competence & competent, capable & 4 
    & \cite{cox2022does}\cmt{[27]}\cite{kraus2020effects}\cmt{[64]}\cite{lee2019s}\cmt{[55]}\cite{westerman2019believe}\cmt{[9]}

@@

\multirow{3}{*}{\parbox{0.16\textwidth}{Perception of Social Connection with Agent}}
    & Conversation Tone & appropriate, expressive, empathetic, persuasive & 13 
    & \cite{chan2021kinvoices}\cmt{[74]}\cite{cox2022does}\cmt{[27]}\cite{cuadra2021my}\cmt{[67]}\cite{daher2020empathic}\cmt{[58]}\cite{diederich2019emulating}\cmt{[25]}\cite{dubiel2020persuasive}\cmt{[60]}\cite{hu2021enhancing}\cmt{[56]}\cite{jestin2022effects}\cmt{[81]}\cite{misu2011toward}\cmt{[83]}\cite{tolmeijer2021female}\cmt{[62]}\cite{yang2021effect}\cmt{[72]}\cite{yang2017perceived}\cmt{[44]}\cite{zhu2022effects}\cmt{[26]}

@@

    & Human-likeness & human-like, natural, artificial, machine-like & 20

@@

    & Personality Traits & friendly, kind, warm, creepy, likeable, polite & 27
    & \cite{andrews2012system}\cmt{[38]}\cite{ashktorab2019resilient}\cmt{[88]}\cite{ceha2022expressive}\cmt{[77]}\cite{ceha2021can}\cmt{[57]}\cite{chan2021kinvoices}\cmt{[74]}\cite{cox2022does}\cmt{[27]}\cite{cuadra2021my}\cmt{[67]}\cite{haas2022keep}\cmt{[78]}\cite{habler2019effects}\cmt{[63]}\cite{healey2013relating}\cmt{[39]}\cite{hu2022polite}\cmt{[76]}\cite{huiyang2022improving}\cmt{[17]}\cite{jeong2019exploring}\cmt{[10]}\cite{jestin2022effects}\cmt{[81]}\cite{khooshabeh2011does}\cmt{[37]}
    \cite{kim2019comparing}\cmt{[89]}\cite{lee2019s}\cmt{[55]}\cite{lee2020hear}\cmt{[23]}\cite{linnemann2018can}\cmt{[15]}\cite{miehle2018exploring}\cmt{[51]}\cite{moilanen2022measuring}\cmt{[82]}\cite{ouchi2019should}\cmt{[59]}\cite{tolmeijer2021female}\cmt{[62]}\cite{volkel2021manipulating}\cmt{[68]}\cite{volkel2022user}\cmt{[75]}\cite{wester2015artificial}\cmt{[14]}\cite{zhu2022effects}\cmt{[26]} 

@@

\begin{table*}[h]
\resizebox{\textwidth}{!}{
\renewcommand*{\arraystretch}{1.4}
\begin{tabular}{@{}p{0.17\textwidth} | p{0.14\textwidth} | p{0.25\textwidth} | >{\centering}p{0.09\textwidth} | p{0.34\textwidth} @{}}
\Xhline{1.2pt}
\textbf{Element Category} & \textbf{Element} & \textbf{Examples} & \textbf{No. Papers} & \textbf{Papers}
\\ \Xhline{1.2pt}
\multirow{4}{*}{Linguistic Content}    & Affective Language & emotional expressions, sentiment-adaptive responses & 11
& \cite{daher2020empathic}\cmt{[58]}\cite{diederich2019emulating}\cmt{[25]}\cite{healey2013relating}\cmt{[39]}\cite{hu2022polite}\cmt{[76]}\cite{lee2019s}\cmt{[55]}\cite{lubis2019positive}\cmt{[43]}\cite{moilanen2022measuring}\cmt{[82]}\cite{seeger2021chatbots}\cmt{[35]}\cite{volkel2022user}\cmt{[75]}\cite{yang2017perceived}\cmt{[44]}\cite{zhu2022effects}\cmt{[26]}
\\ \cline{2-5}
& Social Talk & self-disclosure, social dialogue & 7
& \cite{andrews2012system}\cmt{[38]}\cite{lee2020hear}\cmt{[23]}\cite{lubold2016effects}\cmt{[86]}\cite{moilanen2022measuring}\cmt{[82]}\cite{roy2021users}\cmt{[71]}\cite{volkel2021manipulating}\cmt{[68]}\cite{volkel2022user}\cmt{[75]}
\\ \cline{2-5}
& Humour & humorous content, jokes & 5
& \cite{ceha2021can}\cmt{[57]}\cite{go2021conversational}\cmt{[80]}\cite{khooshabeh2011does}\cmt{[37]}\cite{miyamoto2017improving}\cmt{[46]}\cite{volkel2021manipulating}\cmt{[68]}
\\ \cline{2-5}
& Agent-Initiated Content & proactive dialogue, conversation repair & 4
& \cite{ashktorab2019resilient}\cmt{[88]}\cite{cuadra2021my}\cmt{[67]}\cite{kraus2020effects}\cmt{[64]}\cite{xiao2021let}\cmt{[73]}
\\ \Xhline{1.2pt}
\multirow{4}{*}{Linguistic Style} & Formality & formal vs. casual, honorific expressions & 9
& \cite{cox2022does}\cmt{[27]}\cite{elsholz2019exploring}\cmt{[61]}\cite{habler2019effects}\cmt{[63]}\cite{jestin2022effects}\cmt{[81]}\cite{kim2019comparing}\cmt{[89]}\cite{ma2022ask}\cmt{[29]}\cite{moilanen2022measuring}\cmt{[82]}\cite{ouchi2019should}\cmt{[59]}\cite{volkel2022user}\cmt{[75]}
\\ \cline{2-5}
& Alignment & content matching, lexical alignment, agreeableness & 7
& \cite{healey2013relating}\cmt{[39]}\cite{hoegen2019end}\cmt{[31]}\cite{huiyang2022improving}\cmt{[17]}\cite{linnemann2018can}\cmt{[15]}\cite{spillner2021talk}\cmt{[18]}\cite{volkel2021examining}\cmt{[69]}\cite{volkel2021manipulating}\cmt{[68]}
\\ \cline{2-5}
& Elaborateness & elaborate, concise, sentence structure & 6
& \cite{haas2022keep}\cmt{[78]}\cite{miehle2018exploring}\cmt{[51]}\cite{moilanen2022measuring}\cmt{[82]}\cite{roy2021users}\cmt{[71]}\cite{volkel2021manipulating}\cmt{[68]}\cite{volkel2022user}\cmt{[75]}
\\ \cline{2-5}
& Disfluency & fillers, interjections, repetitions & 6
& \cite{ceha2022expressive}\cmt{[77]}\cite{hu2021enhancing}\cmt{[56]}\cite{jeong2019exploring}\cmt{[10]}\cite{niewiadomski2013laugh}\cmt{[85]}\cite{wester2015artificial}\cmt{[14]}\cite{yang2021effect}\cmt{[72]}
\\ \Xhline{1.2pt}
\multirow{2}{*}{Non-Linguistic Format} & Prosodic Cue & pitch, intonation, speech rate, spoken accent & 12
& \cite{chan2021kinvoices}\cmt{[74]}\cite{choi2020nobody}\cmt{[54]}\cite{dubiel2020persuasive}\cmt{[60]}\cite{feijoo2021effects}\cmt{[70]}\cite{habler2019effects}\cmt{[63]}\cite{hu2021enhancing}\cmt{[56]}\cite{jestin2022effects}\cmt{[81]}\cite{kim2020can}\cmt{[24]}\cite{lubold2016effects}\cmt{[86]}\cite{misu2011toward}\cmt{[83]}\cite{tolmeijer2021female}\cmt{[62]}\cite{zhu2022effects}\cmt{[26]}
\\ \cline{2-5}
& Text Variation & typo, capitalization, emoticons, response delay & 9
& \cite{fadhil2018effect}\cmt{[52]}\cite{gnewuch2018faster}\cmt{[19]}\cite{gnewuch2018chatbot}\cmt{[21]}\cite{gnewuch2022opposing}\cmt{[20]}\cite{kim2019comparing}\cmt{[89]}\cite{seeger2021chatbots}\cmt{[35]}\cite{volkel2022user}\cmt{[75]}\cite{westerman2019believe}\cmt{[9]}\cite{wilhelm2022keep}\cmt{[28]}
\\ \Xhline{1.2pt}
\end{tabular}%
}
\caption{Conversation architecture elements}
\label{tab:cues}
\end{table*}

>>>>

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Wei, Kim, Kuzminykh} %It will be automatically anonymized through the document class property anonymous

@@

\multirow{3}{*}{\parbox{0.17\textwidth}{Perception of \newline Interaction with Agent}}
    & Usability & accuracy, ease of use, \newline efficiency, helpfulness & 23 

@@

    & Engagement & enjoyment, annoyance, \newline desirable, intention to use  & 28

@@

    & Competence & competent, appropriate & 6 
    & \cite{cox2022does}\cmt{[27]}\cite{kraus2020effects}\cmt{[64]}\cite{jestin2022effects}\cmt{[81]}\cite{lee2019s}\cmt{[55]}\cite{misu2011toward}\cmt{[83]}\cite{westerman2019believe}\cmt{[9]}

@@

\multirow{3}{*}{\parbox{0.16\textwidth}{Perception of \newline Sociability with Agent}}
    & Conversation Tone & friendly, warm, empathetic, persuasive & 20 
    & \cite{ashktorab2019resilient}\cmt{[88]}\cite{chan2021kinvoices}\cmt{[74]}\cite{cox2022does}\cmt{[27]}\cite{cuadra2021my}\cmt{[67]}\cite{daher2020empathic}\cmt{[58]}\cite{diederich2019emulating}\cmt{[25]}\cite{dubiel2020persuasive}\cmt{[60]}\cite{healey2013relating}\cmt{[39]}\cite{hu2021enhancing}\cmt{[56]}\cite{hu2022polite}\cmt{[76]}\cite{jestin2022effects}\cmt{[81]}\cite{khooshabeh2011does}\cmt{[37]}\cite{kim2019comparing}\cmt{[89]}\cite{lee2019s}\cmt{[55]}\cite{miehle2018exploring}\cmt{[51]}
    \cite{moilanen2022measuring}\cmt{[82]}\cite{tolmeijer2021female}\cmt{[62]}\cite{yang2021effect}\cmt{[72]}\cite{yang2017perceived}\cmt{[44]}\cite{zhu2022effects}\cmt{[26]}

@@

    & Human-likeness & human-like, natural, \newline artificial, machine-like & 20

@@

    & Personality Traits & likeable, kind, witty, funny, creepy & 20
    & \cite{andrews2012system}\cmt{[38]}\cite{ceha2022expressive}\cmt{[77]}\cite{ceha2021can}\cmt{[57]}\cite{chan2021kinvoices}\cmt{[74]}\cite{cuadra2021my}\cmt{[67]}\cite{haas2022keep}\cmt{[78]}\cite{habler2019effects}\cmt{[63]}\cite{healey2013relating}\cmt{[39]}\cite{hu2022polite}\cmt{[76]}\cite{huiyang2022improving}\cmt{[17]}\cite{jeong2019exploring}\cmt{[10]}\cite{kim2019comparing}\cmt{[89]}\cite{lee2020hear}\cmt{[23]}\cite{linnemann2018can}\cmt{[15]}\cite{miehle2018exploring}\cmt{[51]}
    \cite{ouchi2019should}\cmt{[59]}\cite{volkel2021manipulating}\cmt{[68]}\cite{volkel2022user}\cmt{[75]}\cite{wester2015artificial}\cmt{[14]}\cite{zhu2022effects}\cmt{[26]} 

@@

\begin{table*}[h]
\resizebox{\textwidth}{!}{
\renewcommand*{\arraystretch}{1.4}
\begin{tabular}{@{}p{0.17\textwidth} | p{0.14\textwidth} | p{0.25\textwidth} | >{\centering}p{0.09\textwidth} | p{0.34\textwidth} @{}}
\Xhline{1.2pt}
\multicolumn{2}{l|}{\textbf{Conversation Architecture Elements}} & \textbf{Examples} & \textbf{No. Papers} & \textbf{Papers}
\\ \Xhline{1.2pt}
\multirow{3}{*}{Dialog Strategy} & Social Dialog & social talk, self-disclosure  & 7
& \cite{andrews2012system}\cmt{[38]}\cite{lee2020hear}\cmt{[23]}\cite{lubold2016effects}\cmt{[86]}\cite{moilanen2022measuring}\cmt{[82]}\cite{roy2021users}\cmt{[71]}\cite{volkel2021manipulating}\cmt{[68]}\cite{volkel2022user}\cmt{[75]}
\\ \cline{2-5}

& Initiative & proactive dialogue, conversation \newline repair & 4
& \cite{ashktorab2019resilient}\cmt{[88]}\cite{cuadra2021my}\cmt{[67]}\cite{kraus2020effects}\cmt{[64]}\cite{xiao2021let}\cmt{[73]}
\\ \cline{2-5}

& Response Delay & response delay, typing indicator & 4
& \cite{gnewuch2018faster}\cmt{[19]}\cite{gnewuch2018chatbot}\cmt{[21]}\cite{gnewuch2022opposing}\cmt{[20]}\cite{seeger2021chatbots}\cmt{[35]}
\\ \Xhline{1.2pt}

\multirow{2}{*}{Content Affectiveness} & Affective Language & emotional expressions, sentiment-adaptive responses & 11
& \cite{daher2020empathic}\cmt{[58]}\cite{diederich2019emulating}\cmt{[25]}\cite{healey2013relating}\cmt{[39]}\cite{hu2022polite}\cmt{[76]}\cite{lee2019s}\cmt{[55]}\cite{lubis2019positive}\cmt{[43]}\cite{moilanen2022measuring}\cmt{[82]}\cite{seeger2021chatbots}\cmt{[35]}\cite{volkel2022user}\cmt{[75]}\cite{yang2017perceived}\cmt{[44]}\cite{zhu2022effects}\cmt{[26]}
\\ \cline{2-5}

& Humour & humorous content, jokes & 5
& \cite{ceha2021can}\cmt{[57]}\cite{go2021conversational}\cmt{[80]}\cite{khooshabeh2011does}\cmt{[37]}\cite{miyamoto2017improving}\cmt{[46]}\cite{volkel2021manipulating}\cmt{[68]}
\\ \Xhline{1.2pt}

\multirow{3}{*}{Content Style} & Formality & formal vs. casual, honorific \newline expressions & 9
& \cite{cox2022does}\cmt{[27]}\cite{elsholz2019exploring}\cmt{[61]}\cite{habler2019effects}\cmt{[63]}\cite{jestin2022effects}\cmt{[81]}\cite{kim2019comparing}\cmt{[89]}\cite{ma2022ask}\cmt{[29]}\cite{moilanen2022measuring}\cmt{[82]}\cite{ouchi2019should}\cmt{[59]}\cite{volkel2022user}\cmt{[75]}
\\ \cline{2-5}

& Alignment & content matching, lexical \newline alignment, agreeableness & 7
& \cite{healey2013relating}\cmt{[39]}\cite{hoegen2019end}\cmt{[31]}\cite{huiyang2022improving}\cmt{[17]}\cite{linnemann2018can}\cmt{[15]}\cite{spillner2021talk}\cmt{[18]}\cite{volkel2021examining}\cmt{[69]}\cite{volkel2021manipulating}\cmt{[68]}
\\ \cline{2-5}

& Elaborateness & elaborate, concise, sentence \newline structure & 6
& \cite{haas2022keep}\cmt{[78]}\cite{miehle2018exploring}\cmt{[51]}\cite{moilanen2022measuring}\cmt{[82]}\cite{roy2021users}\cmt{[71]}\cite{volkel2021manipulating}\cmt{[68]}\cite{volkel2022user}\cmt{[75]}
\\ \Xhline{1.2pt}

\multirow{3}{*}{Speech Format} & Prosody & pitch, intonation, speech rate, \newline spoken accent & 12
& \cite{chan2021kinvoices}\cmt{[74]}\cite{choi2020nobody}\cmt{[54]}\cite{dubiel2020persuasive}\cmt{[60]}\cite{feijoo2021effects}\cmt{[70]}\cite{habler2019effects}\cmt{[63]}\cite{hu2021enhancing}\cmt{[56]}\cite{jestin2022effects}\cmt{[81]}\cite{kim2020can}\cmt{[24]}\cite{lubold2016effects}\cmt{[86]}\cite{misu2011toward}\cmt{[83]}\cite{tolmeijer2021female}\cmt{[62]}\cite{zhu2022effects}\cmt{[26]}
\\ \cline{2-5}

& Disfluency & fillers, interjections, repetitions, \newline typos & 7
& \cite{ceha2022expressive}\cmt{[77]}\cite{hu2021enhancing}\cmt{[56]}\cite{jeong2019exploring}\cmt{[10]}\cite{niewiadomski2013laugh}\cmt{[85]}\cite{wester2015artificial}\cmt{[14]}\cite{westerman2019believe}\cmt{[9]}\cite{yang2021effect}\cmt{[72]}
\\ \cline{2-5}

& Text Formatting & capitalization, emoticons & 6
& \cite{fadhil2018effect}\cmt{[52]}\cite{kim2019comparing}\cmt{[89]}\cite{seeger2021chatbots}\cmt{[35]}\cite{volkel2022user}\cmt{[75]}\cite{westerman2019believe}\cmt{[9]}\cite{wilhelm2022keep}\cmt{[28]}
\\ \Xhline{1.2pt}
\end{tabular}%
}
\caption{Conversation architecture elements}
\label{tab:cues}
\end{table*}
